---
title: "When personalization deters delegation to algorithms --- Replication analysis"
author: "Lara Suraci & Irenaeus Wolff"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
    toc: true
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
header-includes:
- \usepackage{fvextra}
- \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,breakanywhere,commandchars=\\\{\}}
---

<!--
============================================================================
REPLICATION PACKAGE -- "When personalization deters delegation to algorithms:
Evidence from a preregistered experiment" (Suraci & Wolff)

This script reproduces every figure and the in-text statistics reported in the
paper. It is organised to follow the structure of the Results section.

Required raw data files (expected in the same folder as this script):
  * 260206_1528_537_lotteryChoices.xls     (decision-level data, tab-separated)
  * allSubjects.csv                        (participant-level data)
  * prolific_demographic_export_approved.csv (Prolific demographics / timing)

Figures produced (file names match the \includegraphics calls in AIms2.tex):
  * Figure1_AverageDelegation.pdf          (Fig. 1 -- main text)
  * Figure3_ECDF.pdf                       (Fig. 3 -- main text)
  * ReasonsByTrmtAndNeverDelegator2.png    (Fig. 4 -- main text)
  * CoefPlotBadAbsurd.pdf                  (Fig. A.1 -- appendix)
  * ExpectedBadAndAbsurdChoicesByTrmt.pdf  (Fig. A.2 -- appendix)

Note: Figure_ExampleScreens.pdf is composed of experiment screenshots and is
not generated here.
============================================================================
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = FALSE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

```{r load-packages}
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(Exact)     # Boschloo exact test
library(fixest)    # LPM with fixed effects + clustered SE
library(lmtest)    # robust coefficient tests
library(sandwich)  # HC2 standard errors
library(texreg)    # regression tables
```

```{r housekeeping, echo=FALSE}
rm(list = ls())
```

\newpage

# Data import and preparation

## Shared plotting constants

To keep figures visually consistent we fix the treatment colour palette once.

```{r plot-constants}
# Treatment colours used throughout (Pref = preference-based, EV = optimisation)
col_pref <- "#4E79A7"
col_ev   <- "#D55E00"
algo_cols <- c("Pref" = col_pref, "EV" = col_ev)
```

## Load raw data

Paths are relative to this script's location (replication-package layout).

```{r load-data}
choices <- read.delim("260206_1528_537_lotteryChoices.xls")
subjects <- read.csv("allSubjects.csv", header = TRUE)
prolificData <- read.csv("prolific_demographic_export_approved.csv")

choicesP1 <- subset(choices, Part == 1)
```

## Preregistered exclusions

We apply the preregistered exclusion rules: failed bot/slider check at first
attempt, failed comprehension checks after the second attempt, failure to
confirm careful participation (`Reliable < 6`), disclosed AI usage, incomplete
data (`Delegated < 0`), and the log-time "too fast" rule.

```{r exclusions}
# "Too fast" flag based on the preregistered log-completion-time rule
prolificData$logCompleTime <- log(prolificData$Time.taken)
prolificData$tooFast <- 1 * (
  prolificData$logCompleTime <
    mean(log(prolificData$Time.taken)) - 2 * sd(log(prolificData$Time.taken))
)

keep_ids <- prolificData$Participant.id[prolificData$tooFast == 0]

choices <- subset(
  choices,
  BotCheck1attempt == 1 &
    Part1CQsPassedOnAttempt <= 2 &
    Part2CQsPassedOnAttempt <= 2 &
    Reliable >= 6 &
    AIUsageInStudy == 0 &
    Delegated >= 0 &
    ProlificID %in% keep_ids
)

subjects  <- subset(subjects,  ProlificID %in% choices$ProlificID)
choicesP1 <- subset(choicesP1, ProlificID %in% choices$ProlificID)

# Final analysis sample (paper: N = 233; EV = 118, Pref = 115)
cat("Final sample size per condition (1 = EV, 0 = Preference-based):\n")
table(subjects$EVAlgorithm)

# Convenience label for the algorithm condition
choices <- choices %>%
  mutate(Algorithm = ifelse(EVAlgorithm == 1, "EV", "Pref"))
```

## Build participant-level Part-1 preference measures

These variables (revealed risk preferences from Part 1, per-task confidence,
and post-task expectations) are merged onto the decision-level `choices` data
for use in the regressions and the appendix coefficient plot.

```{r part1-measures}
# EV-maximising choice in each Part-1 task
choicesP1 <- choicesP1 %>%
  mutate(EVChoice = case_when(
    grepl("^P", ChoiceType) & PBetChosen == 1 ~ 1,
    grepl("^D", ChoiceType) & PBetChosen == 0 ~ 1,
    TRUE ~ 0
  ))

# "True" P-bet choice (less-risky choice when it was NOT EV-maximising)
choicesP1 <- choicesP1 %>%
  mutate(truePChoice = case_when(
    grepl("^D", ChoiceType) & PBetChosen == 1 ~ 1,
    TRUE ~ 0
  ))

# Participant-level Part-1 summaries
p1_summaries <- choicesP1 %>%
  group_by(ProlificID) %>%
  summarise(
    EVChoices    = sum(EVChoice),
    truePChoices = sum(truePChoice),
    PBet_count   = sum(PBetChosen),
    .groups = "drop"
  )

choices <- choices %>%
  left_join(p1_summaries, by = "ProlificID") %>%
  # per-task Part-1 confidence (matched on participant x option set)
  left_join(
    choicesP1 %>% select(ProlificID, OptionsetID, ConfidenceP1 = Confidence),
    by = c("ProlificID", "OptionsetID")
  ) %>%
  # participant-level demographics and post-task expectations
  left_join(
    subjects %>% select(
      ProlificID, Female, Age, Education, HHIncome,
      ProbOfBadComputerChoices, ProbOfAbsurdComputerChoices, ComputerRiskTaking
    ),
    by = "ProlificID"
  )
```

\newpage

# Descriptive overview

## Choice quality and payoffs

Choice quality among non-delegated decisions and the payoff advantage of
delegation (paper, Section "Primary test: first-contact delegation").

```{r descriptives-choicequality}
# Share of non-dominated self-chosen options, by condition
choices %>%
  filter(Part == 2) %>%
  group_by(Algorithm) %>%
  summarise(
    avNonDom = mean(Nondominated[Delegated == 0 & Option > 0], na.rm = TRUE),
    .groups = "drop"
  )

# Average payoff: delegated vs. non-delegated decisions
# (paper: 243.49 points delegated vs. 195.18 points non-delegated, ~25% higher)
choices %>%
  filter(Part == 2) %>%
  group_by(Delegated) %>%
  summarise(meanPayoff = mean(RelevantPayoff, na.rm = TRUE), .groups = "drop")
```

## Round-level delegation profile

```{r round-level-table}
choicesAv <- choices %>%
  filter(Part == 2) %>%
  group_by(Algorithm, WithinPartDecisionNumber) %>%
  summarise(
    avDeleg  = mean(Delegated, na.rm = TRUE),
    avNonDom = mean(Nondominated[Delegated == 0 & Option > 0], na.rm = TRUE),
    .groups = "drop"
  )
choicesAv
```

## Figure 1 --- Average delegation over decisions

```{r figure1-average-delegation, fig.width=7, fig.height=4.5}
fig1 <- choicesAv %>%
  mutate(Algorithm = factor(Algorithm, levels = c("EV", "Pref"),
                            labels = c("EV", "Preference-based"))) %>%
  ggplot(aes(x = WithinPartDecisionNumber, y = avDeleg,
             color = Algorithm, group = Algorithm)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = 1:8) +
  scale_color_manual(values = c("EV" = col_ev, "Preference-based" = col_pref)) +
  labs(x = "Decision number (Part 2)",
       y = "Average delegation",
       color = "Algorithm") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")

fig1
ggsave("Figure1_AverageDelegation.pdf", fig1, width = 7, height = 4.5)
```

\newpage

# Primary test 1 --- First-contact delegation (round 1)

Preregistered one-sided Boschloo exact test on round-1 delegation
(paper: 66.9% vs. 55.8%, p = 0.0413).

```{r boschloo-round1}
tab <- table(
  Delegated   = choices$Delegated[choices$Part == 2 & choices$WithinPartDecisionNumber == 1],
  EVAlgorithm = choices$EVAlgorithm[choices$Part == 2 & choices$WithinPartDecisionNumber == 1]
)

# Delegation share by condition (column 2 = EV)
prop.table(tab, margin = 2)

# One-sided Boschloo exact test (H1: EV > Pref)
exact.test(tab, method = "Boschloo", alternative = "greater", to.plot = FALSE)$p.value
```

\newpage

# Primary test 2 --- Delegation across repeated decisions

## Participant-level Wilcoxon--Mann--Whitney test

Preregistered one-sided WMW test on participant-average delegation
(paper: W = 6428, p = 0.2425).

```{r wmw-test}
delegAv <- choices %>%
  filter(Part == 2) %>%
  group_by(Algorithm, ProlificID) %>%
  summarise(avDeleg = mean(Delegated, na.rm = TRUE), .groups = "drop") %>%
  mutate(Algorithm = factor(Algorithm, levels = c("Pref", "EV")))

wilcox.test(avDeleg ~ Algorithm, data = delegAv, alternative = "less")
```

## Figure 3 --- ECDF of participant-level average delegation

```{r figure3-ecdf, fig.width=7, fig.height=4.5}
fig3 <- delegAv %>%
  mutate(Algorithm = factor(Algorithm, levels = c("EV", "Pref"),
                            labels = c("EV", "Preference-based"))) %>%
  ggplot(aes(x = avDeleg, color = Algorithm)) +
  stat_ecdf(linewidth = 1) +
  scale_color_manual(values = c("EV" = col_ev, "Preference-based" = col_pref)) +
  labs(x = "Participant-level average delegation",
       y = "Cumulative probability",
       color = "Algorithm") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")

fig3
ggsave("Figure3_ECDF.pdf", fig3, width = 7, height = 4.5)
```

## Preregistered regression specification

LPM with decision-set fixed effects and participant-clustered SEs. We first
construct the feedback-history regressor (running share of zero outcomes among
prior delegated choices).

```{r feedback-history}
choices <- choices %>%
  arrange(ProlificID, WithinPartDecisionNumber) %>%
  group_by(ProlificID) %>%
  mutate(
    delegated_round = (Delegated == 1),
    zero_payoff     = (RelevantPayoff == 0 & Delegated == 1),
    cum_delegated   = lag(cumsum(delegated_round), default = 0),
    cum_zero        = lag(cumsum(zero_payoff), default = 0),
    zero_rate       = ifelse(cum_delegated > 0, cum_zero / cum_delegated, NA)
  ) %>%
  ungroup()
```

```{r preregistered-regression}
# Preregistered model: treatment x (period + feedback history)
delRegPreReg <- feols(
  Delegated ~ EVAlgorithm * (WithinPartDecisionNumber + zero_rate) | OptionsetID,
  cluster = ~ProlificID,
  data = subset(choices, Part == 2)
)
summary(delRegPreReg)
```

## Supporting regressions

Plain treatment effect (all rounds; excluding round 1; round 1 only). The
round-1-only LPM reproduces the Boschloo result
($\beta = 0.113$, one-sided $p = 0.041$).

```{r supporting-regressions}
# All rounds, treatment only
delRegPlainTrmt <- feols(
  Delegated ~ EVAlgorithm | OptionsetID,
  cluster = ~ProlificID,
  data = subset(choices, Part == 2)
)

# Excluding round 1
delRegPlainTrmtNoP1 <- feols(
  Delegated ~ EVAlgorithm | OptionsetID,
  cluster = ~ProlificID,
  data = subset(choices, Part == 2 & WithinPartDecisionNumber > 1)
)

# Round 1 only (mirrors the Boschloo test)
delRegP1 <- feols(
  Delegated ~ EVAlgorithm | OptionsetID,
  cluster = ~ProlificID,
  data = subset(choices, Part == 2 & WithinPartDecisionNumber == 1)
)

screenreg(
  list(delRegPlainTrmt, delRegPlainTrmtNoP1, delRegP1),
  custom.model.names = c("All rounds", "Rounds 2-8", "Round 1 only"),
  caption = "Treatment effect on delegation (LPM, task FE, clustered SE)"
)
```

\newpage

# Categorical non-delegation

We classify participants as never-, sometimes-, or always-delegators and
compare never-delegation across conditions (paper: 15.7% Pref vs. 3.4% EV).

```{r delegator-types}
choices <- choices %>%
  group_by(ProlificID) %>%
  mutate(sumDelegations = sum(Delegated)) %>%
  ungroup() %>%
  left_join(
    prolificData %>% select(Participant.id, TimeInSeconds = Time.taken),
    by = c("ProlificID" = "Participant.id")
  )

analysis_data <- choices %>%
  group_by(ProlificID) %>%
  summarise(
    EVAlgorithm                 = first(EVAlgorithm),
    sumDelegations              = first(sumDelegations),
    Female                      = first(Female),
    Age                         = first(Age),
    HHIncome                    = first(HHIncome),
    Education                   = first(Education),
    ProbOfBadComputerChoices    = first(ProbOfBadComputerChoices),
    ProbOfAbsurdComputerChoices = first(ProbOfAbsurdComputerChoices),
    ComputerRiskTaking          = first(ComputerRiskTaking),
    PBetChoices                 = first(PBet_count),
    TimeInSeconds               = first(TimeInSeconds),
    .groups = "drop"
  ) %>%
  mutate(
    DelegationGroup = ifelse(sumDelegations == 0, "Never", "AtLeastOnce"),
    NonDelegator    = 1 * (sumDelegations == 0),
    DelegatorType   = factor(
      if_else(sumDelegations == 0, "NeverDelegator",
              if_else(sumDelegations == 8, "AlwaysDelegator", "SometimesDelegator")),
      levels = c("NeverDelegator", "SometimesDelegator", "AlwaysDelegator")
    )
  ) %>%
  left_join(
    subjects %>% select(ProlificID, RiskTaking, FinMarketExperience,
                        FinanciallyResponsible, AIXperience),
    by = "ProlificID"
  )

# Distribution of delegator types by condition (column 2 = EV)
prop.table(table(analysis_data$DelegatorType, analysis_data$EVAlgorithm), margin = 2)
```

## Comparisons of never-delegators vs. delegators

The following tests reproduce the in-text comparisons (age, income, education,
time, gender, risk-taking, reliability, AI experience, Part-1 satisfaction).

```{r nondelegator-tests}
# Age
t.test(Age ~ DelegationGroup, data = analysis_data)

# Time in experiment
t.test(TimeInSeconds ~ NonDelegator, data = analysis_data)

# Household income (categorical)
chisq.test(table(analysis_data$DelegationGroup, analysis_data$HHIncome))

# Education (categorical)
chisq.test(table(analysis_data$DelegationGroup, analysis_data$Education))

# Gender: never-delegation rate among males vs. females (Boschloo)
female_table <- table(
  DelegationGroup = analysis_data$DelegationGroup,
  Female          = analysis_data$Female
)
exact.test(female_table[, c(2, 4)], method = "Boschloo", to.plot = FALSE)$p.value

# Self-reported financial risk-taking, reliability, AI experience
t.test(RiskTaking  ~ NonDelegator, data = analysis_data)
t.test(FinanciallyResponsible ~ NonDelegator, data = analysis_data)
t.test(AIXperience ~ NonDelegator, data = analysis_data)
```

```{r nondelegator-satisfaction-pref}
# Within the preference-based condition: Part-1 satisfaction by delegation type
NeverDelegate <- choices %>%
  select(ProlificID, sumDelegations) %>%
  distinct() %>%
  mutate(sumDeleg0 = (sumDelegations == 0))

subjects2 <- subjects %>% left_join(NeverDelegate, by = "ProlificID")

subjects2 %>%
  filter(EVAlgorithm == 0) %>%
  group_by(sumDeleg0) %>%
  summarise(SatisfiedPart1 = mean(SatisfiedPart1, na.rm = TRUE),
            n = n(), .groups = "drop")
```

\newpage

# Responses to feedback (win-stay / lose-shift)

Win-stay/lose-shift model on rounds 2--8 (paper, Section "Responses to
feedback").

```{r win-stay-lose-shift}
choices <- choices %>%
  group_by(ProlificID) %>%
  arrange(WithinPartDecisionNumber, .by_group = TRUE) %>%
  mutate(
    lagDelegated          = lag(Delegated, 1),
    lagRelevantPayoff     = lag(RelevantPayoff, 1),
    lagSatisfactionRating = lag(SatisfactionRating, 1)
  ) %>%
  ungroup()

choices$stay <- 1 * (choices$Delegated == choices$lagDelegated)

# "Prior win": the outcome of the PREVIOUS round was strictly positive.
# (lagWin is defined from the lagged payoff, i.e. the genuine win-stay/lose-shift
# notion of reacting to the previously experienced outcome.)
choices$lagWin <- 1 * (choices$lagRelevantPayoff > 0)

# Pooled model: treatment x prior win
winStayReg <- feols(
  stay ~ EVAlgorithm * lagWin | OptionsetID + WithinPartDecisionNumber,
  cluster = ~ProlificID,
  data = subset(choices, Part == 2 & WithinPartDecisionNumber > 1)
)
summary(winStayReg)
```

Separately by treatment, distinguishing whether the previous round was
delegated (`lagDelegated == 1`, i.e. "the algorithm" produced the outcome) from
a self-made choice. This separates lose-shift after one's own loss from
lose-shift after a delegated (algorithmic) loss.

```{r win-stay-by-treatment}
# EV condition
winStayReg_EV <- feols(
  stay ~ lagDelegated * I(1 - lagWin) | OptionsetID + WithinPartDecisionNumber,
  cluster = ~ProlificID,
  data = subset(choices, EVAlgorithm == 1 & Part == 2 & WithinPartDecisionNumber > 1)
)
summary(winStayReg_EV)

# Preference-based condition
winStayReg_Pref <- feols(
  stay ~ lagDelegated * I(1 - lagWin) | OptionsetID + WithinPartDecisionNumber,
  cluster = ~ProlificID,
  data = subset(choices, EVAlgorithm == 0 & Part == 2 & WithinPartDecisionNumber > 1)
)
summary(winStayReg_Pref)
```

```{r win-stay-rates}
# Average stay rates by condition, prior delegation, and prior win/loss
choices %>%
  filter(Part == 2, WithinPartDecisionNumber > 1) %>%
  group_by(EVAlgorithm, lagDelegated, lagWin) %>%
  summarise(avg_stay_rate = mean(stay, na.rm = TRUE),
            n = n(), .groups = "drop") %>%
  arrange(EVAlgorithm, lagDelegated, lagWin)
```


\newpage

# Exploratory mechanisms

## Coefficient model (interaction with performance expectations)

LPM of individual delegation decisions interacting treatment with participants'
algorithmic performance expectations, controlling for round, per-choice Part-1
confidence, revealed Part-1 risk preferences, and task fixed effects.

```{r mechanism-regression}
delRegWConf <- feols(
  Delegated ~ EVAlgorithm * (WithinPartDecisionNumber + ConfidenceP1 +
    ProbOfAbsurdComputerChoices + ProbOfBadComputerChoices +
    EVChoices + truePChoices) | OptionsetID,
  cluster = ~ProlificID,
  data = subset(choices, Part == 2)
)
summary(delRegWConf)
```

## Figure A.1 --- Condition-specific marginal effects

```{r figureA1-coefplot, fig.width=7, fig.height=4.5}
coefs_wc <- coef(delRegWConf)
vcm_wc   <- vcov(delRegWConf)

# SE of the sum of two coefficients (main effect + interaction)
se_sum <- function(v1, v2) {
  sqrt(vcm_wc[v1, v1] + vcm_wc[v2, v2] + 2 * vcm_wc[v1, v2])
}

coef_spec <- data.frame(
  var_name = c("WithinPartDecisionNumber", "ConfidenceP1",
               "ProbOfBadComputerChoices", "ProbOfAbsurdComputerChoices",
               "EVChoices", "truePChoices"),
  var_label = c("Decision round", "Confidence (Part 1)",
                "Number of non-preferred choices",
                "Number of completely weird choices",
                "EV-maximising choices (Part 1)",
                "P-bet choices when not EV-maximising (Part 1)"),
  stringsAsFactors = FALSE
)

plot_data <- data.frame()
for (i in seq_len(nrow(coef_spec))) {
  var_name  <- coef_spec$var_name[i]
  var_label <- coef_spec$var_label[i]

  pb_est <- coefs_wc[var_name]
  pb_se  <- sqrt(vcm_wc[var_name, var_name])

  int_name <- paste0("EVAlgorithm:", var_name)
  if (int_name %in% names(coefs_wc)) {
    ev_est <- pb_est + coefs_wc[int_name]
    ev_se  <- se_sum(var_name, int_name)
  } else {
    ev_est <- pb_est
    ev_se  <- pb_se
  }

  plot_data <- rbind(plot_data, data.frame(
    Variable  = var_label,
    Condition = c("Preference-based", "EV"),
    Estimate  = c(pb_est, ev_est),
    SE        = c(pb_se, ev_se),
    stringsAsFactors = FALSE
  ))
}

plot_data <- plot_data %>%
  mutate(
    Variable  = factor(Variable, levels = coef_spec$var_label),
    Condition = factor(Condition, levels = c("Preference-based", "EV")),
    CI_lo = Estimate - 1.96 * SE,
    CI_hi = Estimate + 1.96 * SE,
    y_pos = as.numeric(Variable) +
      ifelse(Condition == "Preference-based", 0.15, -0.15)
  )

figA1 <- ggplot(plot_data) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray70") +
  geom_linerange(aes(xmin = CI_lo, xmax = CI_hi, y = y_pos, color = Condition),
                 linewidth = 0.6, lineend = "round") +
  geom_point(aes(x = Estimate, y = y_pos, color = Condition, shape = Condition),
             size = 2.5) +
  scale_y_continuous(breaks = seq_len(nrow(coef_spec)),
                     labels = levels(plot_data$Variable),
                     expand = c(0.15, 0)) +
  scale_color_manual(values = c("Preference-based" = "#E66100", "EV" = "#5D3A9B")) +
  scale_shape_manual(values = c("Preference-based" = 16, "EV" = 17)) +
  labs(x = "Effect on delegation probability", y = NULL,
       color = NULL, shape = NULL,
       caption = paste0("95% confidence intervals. Linear probability model with ",
                        "task fixed effects and participant-clustered SEs.\n",
                        "EV-condition effects are main effect + interaction term.")) +
  theme_bw(base_size = 10) +
  theme(legend.position = "bottom",
        panel.grid.major.y = element_blank(),
        plot.caption = element_text(hjust = 0, size = 9))

figA1
ggsave("CoefPlotBadAbsurd.pdf", figA1, width = 7, height = 4.5)
```

## Figure A.2 --- Expected non-preferred and completely weird choices

```{r figureA2-expectations, fig.width=7, fig.height=4}
figA2 <- subjects %>%
  pivot_longer(cols = c(ProbOfBadComputerChoices, ProbOfAbsurdComputerChoices),
               names_to = "variable", values_to = "value") %>%
  ggplot(aes(x = value, fill = factor(EVAlgorithm))) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ variable, labeller = labeller(variable = c(
    ProbOfBadComputerChoices    = "Non-preferred choices",
    ProbOfAbsurdComputerChoices = "Completely weird choices"
  ))) +
  scale_fill_manual(values = c("0" = "#E66100", "1" = "#5D3A9B"),
                    labels = c("0" = "Preference-based", "1" = "EV"),
                    name = "Condition") +
  labs(x = "Expected number of choices (out of 10)", y = "Density") +
  theme_bw() +
  theme(legend.position = "bottom")

figA2
ggsave("ExpectedBadAndAbsurdChoicesByTrmt.pdf", figA2, width = 7, height = 4)
```

## Treatment comparison of expected algorithmic errors

Formal comparison of expected non-preferred and completely weird choices across
treatments. A LOWER expected number of non-preferred choices under the
preference-based algorithm would indicate that participants believe the
personalized algorithm is doing its job (matching their preferences).

```{r expectations-comparison}
expect_tbl <- subjects %>%
  mutate(Condition = ifelse(EVAlgorithm == 1, "EV", "Preference-based")) %>%
  group_by(Condition) %>%
  summarise(
    n                  = n(),
    mean_nonpreferred  = mean(ProbOfBadComputerChoices, na.rm = TRUE),
    sd_nonpreferred    = sd(ProbOfBadComputerChoices, na.rm = TRUE),
    mean_weird         = mean(ProbOfAbsurdComputerChoices, na.rm = TRUE),
    sd_weird           = sd(ProbOfAbsurdComputerChoices, na.rm = TRUE),
    .groups = "drop"
  )
expect_tbl

# Expected NON-PREFERRED choices: EV vs. Preference-based
t.test(ProbOfBadComputerChoices ~ EVAlgorithm, data = subjects)
wilcox.test(ProbOfBadComputerChoices ~ EVAlgorithm, data = subjects)

# Expected COMPLETELY WEIRD choices: EV vs. Preference-based
t.test(ProbOfAbsurdComputerChoices ~ EVAlgorithm, data = subjects)
wilcox.test(ProbOfAbsurdComputerChoices ~ EVAlgorithm, data = subjects)

# Within the preference-based condition: do never-delegators expect the
# personalized algorithm to make MORE non-preferred choices than delegators?
# (If not, categorical non-delegation is not driven by distrust in the
#  algorithm's ability to capture one's preferences.)
analysis_data %>%
  filter(EVAlgorithm == 0) %>%
  group_by(NonDelegator) %>%
  summarise(
    n                 = n(),
    mean_nonpreferred = mean(ProbOfBadComputerChoices, na.rm = TRUE),
    .groups = "drop"
  )

t.test(ProbOfBadComputerChoices ~ NonDelegator,
       data = subset(analysis_data, EVAlgorithm == 0))
```

\newpage

# Post-task reasons

## Figure 4 --- Reasons by treatment and delegation type

Reason-selection frequencies for the hypothetical additional-task choice, by
treatment (EV vs. preference-based) and delegation type (delegators vs.
never-delegators).

```{r figure4-reasons, fig.width=8, fig.height=4}
subjects2 <- subjects %>% left_join(NeverDelegate, by = "ProlificID")

subjects_long <- subjects2 %>%
  pivot_longer(cols = DontTrust:Other, names_to = "item", values_to = "clicked")

# Group sizes for legend labels
n_labels <- subjects_long %>%
  group_by(sumDeleg0, EVAlgorithm) %>%
  summarise(n = n_distinct(ProlificID), .groups = "drop") %>%
  mutate(
    group = case_when(
      sumDeleg0 == TRUE  & EVAlgorithm == 0 ~ "Prefs--Never-delegators",
      sumDeleg0 == TRUE  & EVAlgorithm == 1 ~ "EV--Never-delegators",
      sumDeleg0 == FALSE & EVAlgorithm == 0 ~ "Prefs--Delegators",
      sumDeleg0 == FALSE & EVAlgorithm == 1 ~ "EV--Delegators"
    ),
    group_label = paste0(group, " (n=", n, ")")
  )

group_colors <- c(
  "Prefs--Delegators"       = "#9dc3e6",
  "Prefs--Never-delegators" = "#1f4e79",
  "EV--Delegators"          = "#f4b183",
  "EV--Never-delegators"    = "#833c00"
)
label_map <- setNames(n_labels$group_label, n_labels$group)

click_rates_both <- subjects_long %>%
  group_by(sumDeleg0, EVAlgorithm, item) %>%
  summarise(rate = mean(clicked, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    group = case_when(
      sumDeleg0 == TRUE  & EVAlgorithm == 0 ~ "Prefs--Never-delegators",
      sumDeleg0 == TRUE  & EVAlgorithm == 1 ~ "EV--Never-delegators",
      sumDeleg0 == FALSE & EVAlgorithm == 0 ~ "Prefs--Delegators",
      sumDeleg0 == FALSE & EVAlgorithm == 1 ~ "EV--Delegators"
    ),
    group = factor(group, levels = names(group_colors))
  )

fig4 <- ggplot(click_rates_both, aes(x = item, y = rate, fill = group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.7) +
  scale_fill_manual(values = group_colors, labels = label_map,
                    breaks = names(group_colors), name = "Condition") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "Item", y = "Click rate") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right")

fig4
ggsave("ReasonsByTrmtAndNeverDelegator2.png", fig4, width = 8, height = 4, dpi = 300)
```

## Aggregate reason patterns

In-text figures for performance, distrust, and responsibility motives by
delegation type (delegators vs. never-delegators).

```{r reasons-aggregate}
subjects_long %>%
  group_by(sumDeleg0, item) %>%
  summarise(rate = mean(clicked, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = sumDeleg0, values_from = rate) %>%
  rename(Delegators = `FALSE`, NeverDelegators = `TRUE`)
```

## LPM: never-delegation on reasons

```{r lpm-reasons}
reason_vars <- c("DontTrust", "BetterJob", "HowExactly", "MoreInfo",
                 "Responsibility4Choice", "Responsibility4Outcome",
                 "DislikeRelyingOn", "Reactance2Analysis",
                 "PerformanceCuriosity", "Tedious", "Other")

# Bivariate LPMs (one per reason, HC2 SEs)
bivariate_tbl <- do.call(rbind, lapply(reason_vars, function(rv) {
  m  <- lm(as.formula(paste("as.numeric(sumDeleg0) ~", rv)), data = subjects2)
  ct <- coeftest(m, vcov = vcovHC(m, type = "HC2"))
  data.frame(reason = rv,
             coef = round(ct[2, 1], 3), se = round(ct[2, 2], 3),
             t = round(ct[2, 3], 3), p = round(ct[2, 4], 3))
}))
print(bivariate_tbl, row.names = FALSE)

# Joint LPM: full sample + preference-treatment only
f_joint  <- as.formula(paste("as.numeric(sumDeleg0) ~",
                             paste(reason_vars, collapse = " + ")))
lpm_all  <- lm(f_joint, data = subjects2)
lpm_pref <- lm(f_joint, data = subjects2[subjects2$EVAlgorithm == 0, ])

vcov_all  <- vcovHC(lpm_all,  type = "HC2")
vcov_pref <- vcovHC(lpm_pref, type = "HC2")

screenreg(
  list(lpm_all, lpm_pref),
  override.se = list(sqrt(diag(vcov_all)), sqrt(diag(vcov_pref))),
  override.pvalues = list(
    2 * pt(abs(coef(lpm_all)  / sqrt(diag(vcov_all))),
           df = lpm_all$df.residual,  lower.tail = FALSE),
    2 * pt(abs(coef(lpm_pref) / sqrt(diag(vcov_pref))),
           df = lpm_pref$df.residual, lower.tail = FALSE)
  ),
  custom.model.names = c("Full sample", "Pref. treatment"),
  caption = "LPM: never-delegation ~ reasons (HC2 SEs)"
)
```

# Session information

```{r session-info}
sessionInfo()
```
