---
title: "AIms Study: Pre-registered & Exploratory Analysis"
author: "GitHub Copilot"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: '3'
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, fig.width=8, fig.height=5)
library(tidyverse)
library(lubridate)
library(broom)
library(fixest)
```

# 1. Data Loading and Variable Mapping

```{r load-data}
all_subj <- read_csv('allSubjects.csv')
lottery <- read_delim('260206_1528_537_lotteryChoices.xls', delim='\t')
demo <- read_csv('prolific_demographic_export_approved.csv')

# Basic sanity checks
cat('allSubjects rows:', nrow(all_subj), '\n')
cat('lotteryChoices rows:', nrow(lottery), '\n')
cat('demographics rows:', nrow(demo), '\n')
```

## 1.1 Data cleaning and mapping variables

```{r clean-data}
# Core participant-level dataset
participants <- all_subj %>%
  mutate(
    condition = case_when(
      EVAlgorithm == 1 ~ 'EV',
      PrefAlgorithm == 1 ~ 'PREF',
      TRUE ~ NA_character_
    ),
    comp_ok = (CompQ205a == 205 & CompQ10a == 10 & CompQ3a == 3),
    slider_ok = (BotCheck1attempt == 1),
    not_ai = (AIUsageInStudy == 0),
    complete_ok = (LeaveStage == 0)
  ) %>%
  left_join(demo %>% select(`Participant id`, Age, Sex, `Country of residence`, `Employment status`),
            by = c('ProlificID' = 'Participant id'))

# Confirm condition counts
participants %>% count(condition)
```

# 2. Pre-registered Exclusion Filters

```{r exclusion-summary}
initial_n <- nrow(participants)
filtered <- participants %>%
  filter(slider_ok, comp_ok, not_ai, complete_ok)
filtered_n <- nrow(filtered)

exclusion_report <- tibble(
  step = c('initial', 'slider+comprehension+AI+complete'),
  n = c(initial_n, filtered_n)
)
exclusion_report
```

# 3. Delegation Outcomes in Lottery Choices

```{r delegation-aggregation}
# Keep part 2 choices only (decision-taking tasks) and join participant condition
decisions <- lottery %>%
  filter(Part == 2) %>%
  left_join(filtered %>% select(ProlificID, condition, participant_TreatmentNumber = TreatmentNumber),
            by = 'ProlificID') %>%
  mutate(TreatmentNumber = participant_TreatmentNumber,
         delegated = as.integer(Delegated == 1),
         round = WithinPartDecisionNumber) %>%
  select(-participant_TreatmentNumber)

# Per participant round-level delegation
participant_delegate <- decisions %>%
  group_by(ProlificID, TreatmentNumber, condition) %>%
  summarize(
    n_rounds = n(),
    delegated_round1 = delegated[round == 1][1],
    mean_delegation = mean(delegated, na.rm=TRUE),
    .groups='drop'
  )

# Check per-participant number of decisions
participant_delegate %>% summarise(min=n_rounds, max=n_rounds, median=n_rounds)
```

# 4. Primary Pre-Registered Tests

## 4.1 Delegation in Round 1: EV vs PREF

```{r round1-tests}
round1 <- decisions %>%
  filter(round == 1, !is.na(condition)) %>%
  count(condition, delegated) %>%
  pivot_wider(names_from = delegated, values_from = n, values_fill = 0)

# Normalize column names to delegation/no_delegation
colnames(round1) <- sub('^0$', 'no_delegation', colnames(round1))
colnames(round1) <- sub('^1$', 'delegation', colnames(round1))

# If delegated is logical, rename TRUE/FALSE
colnames(round1) <- sub('^FALSE$', 'no_delegation', colnames(round1))
colnames(round1) <- sub('^TRUE$', 'delegation', colnames(round1))

if (!all(c('EV', 'PREF') %in% round1$condition) || nrow(round1) < 2) {
  cat('Round1 cohort does not have both EV and PREF groups or no data to run Fisher test.\n')
  print(round1)
  fisher_result <- NULL
} else if (!all(c('delegation','no_delegation') %in% colnames(round1))) {
  cat('Round1 structure missing delegation/no_delegation columns; check delegated coding.\n')
  print(round1)
  fisher_result <- NULL
} else {
  m <- round1 %>%
    arrange(condition) %>%
    select(delegation, no_delegation) %>%
    as.matrix()

  rownames(m) <- round1 %>% arrange(condition) %>% pull(condition)
  colnames(m) <- c('delegation','no_delegation')

  cat('Round1 contingency table (rows=condition, cols=delegation/no_delegation):\n')
  print(m)

  if (nrow(m) < 2 || ncol(m) < 2) {
    cat('Contingency table is not 2x2; cannot run Fisher test.\n')
    fisher_result <- NULL
  } else {
    fisher_result <- fisher.test(m, alternative='two.sided')
    fisher_result
  }
}
```

## 4.2 Average Delegation Frequency Over Part 2

```{r avg-delegation-tests}
# Use participant-level mean delegation from part2
mean_by_condition <- participant_delegate %>% group_by(condition) %>% summarize(n=n(), mean_delegation=mean(mean_delegation), sd=sd(mean_delegation))
mean_by_condition

condition_levels <- participant_delegate %>% pull(condition) %>% unique() %>% na.omit()
if (length(condition_levels) != 2) {
  cat('Wilcoxon test skipped: condition factor does not have exactly 2 levels.\n')
  cat('condition levels:', paste(condition_levels, collapse=', '), '\n')
  wilcox_result <- NULL
} else {
  wilcox_result <- wilcox.test(mean_delegation ~ condition, data=participant_delegate, exact = FALSE)
  wilcox_result
}
```

## 4.3 Pre-registered regression with clustering

```{r logit-regression}
# Add prior 0 outcomes measure in part2 for each decision
decisions2 <- decisions %>%
  arrange(SessionID, TreatmentNumber, round) %>%
  group_by(SessionID, TreatmentNumber, ProlificID) %>%
  mutate(
    prior_zero = lag(cummean(as.integer(RelevantPayoff == 0)), default=0),
    round_factor = factor(round)
  ) %>%
  ungroup() %>%
  filter(!is.na(condition))

# LPM with participant clustering and decision fixed effects using fixest
if (nrow(decisions2) == 0) {
  cat('Regression skipped: decisions2 has 0 observations after filtering.\n')
  reg <- NULL
} else {
  reg <- feols(delegated ~ condition * round + prior_zero + (1 | OptionsetID), 
               data = decisions2, cluster = ~TreatmentNumber)
  summary(reg)
}
```

# 5. Secondary Pre-registered Hypothesis: Forgiveness after 0 payoff

```{r conditional-switching}
# Coach: compute previous round payoff and current decision to delegate
# derive whether last choice had 0 payoff and whether participant switches from delegation to non-delegation
s <- decisions2 %>%
  group_by(SessionID, TreatmentNumber, ProlificID) %>%
  arrange(round) %>%
  mutate(prev_delegated = lag(delegated),
         prev_zero = lag(RelevantPayoff == 0),
         switched = (prev_delegated == 1 & delegated == 0)) %>%
  ungroup() %>%
  filter(!is.na(prev_delegated))

forgiveness <- s %>%
  filter(prev_zero == TRUE) %>%
  group_by(condition) %>%
  summarize(n=n(), switch_rate=mean(switched, na.rm=TRUE))
forgiveness

# Compare with logistic regression
if (nrow(s) == 0) {
  cat('Switch regression skipped: no observations in forgiveness dataset.\n')
  switch_reg <- NULL
} else {
  switch_reg <- feols(switched ~ condition + prev_zero + condition:prev_zero, data=s, cluster=~TreatmentNumber)
  summary(switch_reg)
}
```

# 6. Exploratory Feature Analysis

```{r feature-exploratory}
# Merge education and demographics
available_cols <- colnames(participants)
need_cols <- c('SessionID','TreatmentNumber','ProlificID','condition')
use_cols <- intersect(c(need_cols,'Age','Female','Education','HHIncome','AIUsageInStudy'), available_cols)
feat <- participants %>% select(all_of(intersect(c('SessionID','TreatmentNumber','ProlificID','condition'), colnames(participants))))
if ('Age' %in% colnames(participants)) feat <- feat %>% mutate(Age = participants$Age)
if ('Female' %in% colnames(participants)) feat <- feat %>% mutate(Female = participants$Female)
if ('Education' %in% colnames(participants)) feat <- feat %>% mutate(Education = participants$Education)
if ('HHIncome' %in% colnames(participants)) feat <- feat %>% mutate(HHIncome = participants$HHIncome)
if ('AIUsageInStudy' %in% colnames(participants)) feat <- feat %>% mutate(AIUsageInStudy = participants$AIUsageInStudy)

if ('Female' %in% colnames(feat)) {
  feat <- feat %>% mutate(Female = if_else(Female == 1, 'Female', 'Not Female'))
}

# Participant aggregate mean delegation
feat <- feat %>% left_join(participant_delegate %>% select(ProlificID, mean_delegation),
                            by = 'ProlificID')

features <- c('condition', intersect(c('Age','Female','Education','HHIncome'), names(feat)))
if (length(features) < 2 || !('mean_delegation' %in% names(feat))) {
  cat('Exploratory feature model skipped: insufficient features or outcome missing.\n')
  feat_model <- NULL
} else {
  if (n_distinct(feat$mean_delegation, na.rm=TRUE) <= 1) {
    cat('Exploratory feature model skipped: dependent variable is constant.\n')
    feat_model <- NULL
  } else {
    formula <- as.formula(paste('mean_delegation ~', paste(features, collapse=' + ')))
    feat_model <- feols(formula, data=feat, na.action='na.omit')
    summary(feat_model)
  }
}
```

# 7. Results Summary

```{r results-summary}
cat('Final pre-registered analysis sample size (participant-level):', nrow(participant_delegate), '\n')
cat('Round1 Fisher p-value:', fisher_result$p.value, '\n')
cat('Mann-Whitney p-value:', wilcox_result$p.value, '\n')
cat('Mean delegation (EV):', mean_by_condition$mean_delegation[mean_by_condition$condition=='EV'], '\n')
cat('Mean delegation (PREF):', mean_by_condition$mean_delegation[mean_by_condition$condition=='PREF'], '\n')
```
