---
title: "When personalization deters delegation to algorithms --- Replication analysis"
author: "Lara Suraci & Irenaeus Wolff"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
  html_document:
    df_print: paged
    toc: true
header-includes:
- \usepackage{fvextra}
- \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,breakanywhere,commandchars=\\\{\}}
---

<!--
============================================================================
REPLICATION PACKAGE -- "When personalization deters delegation to algorithms:
Evidence from a preregistered experiment" (Suraci & Wolff)

This script reproduces every figure and the in-text statistics reported in the
paper. It is organised to follow the structure of the Results section.

Required raw data files (expected in the same folder as this script):
  * 260206_1528_537_lotteryChoices.xls     (decision-level data, tab-separated)
  * allSubjects.csv                        (participant-level data)
  * prolific_demographic_export_approved.csv (Prolific demographics / timing)
  * pilot_data_complete_iw.csv             (earlier, non-preregistered pilot study;
                                             only needed for the retrospective power
                                             analysis near the end of this script)
  * FullTaskTableWithExemplaryOrdering.ods (Part-1/Part-2 payoff matrices; only
                                             needed for the Appendix task-matrix
                                             chunk near the end of this script;
                                             requires the readODS package)

Figures produced (file names match the \includegraphics calls in AIms2.tex):
  * Figure1_AverageDelegation.pdf          (Fig. 1 -- main text)
  * Figure3_ECDF.pdf                       (Fig. 3 -- main text)
  * ReasonsByTrmtAndNeverDelegator2.png    (Fig. 4 -- main text)
  * CoefPlotBadAbsurd.pdf                  (Fig. A.1 -- appendix)
  * ExpectedBadAndAbsurdChoicesByTrmt.pdf  (Fig. A.2 -- appendix)

Note: Figure_ExampleScreens.pdf is composed of experiment screenshots and is
not generated here.
============================================================================
-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = FALSE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

```{r load-packages}
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(Exact)     # Boschloo exact test
library(fixest)    # LPM with fixed effects + clustered SE
library(lmtest)    # robust coefficient tests
library(sandwich)  # HC2 standard errors
library(texreg)    # regression tables
# library(readODS) # only needed for the Appendix task-matrix chunk near the end
```

```{r housekeeping, echo=FALSE}
rm(list = ls())
```

\newpage

# Data import and preparation

## Shared plotting constants

To keep figures visually consistent we fix the treatment colour palette once.

```{r plot-constants}
# Treatment colours used throughout (Pref = preference-based, EV = optimisation)
col_pref <- "#4E79A7"
col_ev   <- "#D55E00"
algo_cols <- c("Pref" = col_pref, "EV" = col_ev)
```

## Load raw data

Paths are relative to this script's location (replication-package layout).

```{r load-data}
choices <- read.delim("260206_1528_537_lotteryChoices.xls")
subjects <- read.csv("allSubjects.csv", header = TRUE)
prolificData <- read.csv("prolific_demographic_export_approved.csv")

choicesP1 <- subset(choices, Part == 1)
```

## Preregistered exclusions

We apply the preregistered exclusion rules: failed bot/slider check at first
attempt, failed comprehension checks after the second attempt, failure to
confirm careful participation (`Reliable < 6`), disclosed AI usage, incomplete
data (`Delegated < 0`), and the log-time "too fast" rule.

```{r exclusions}
# "Too fast" flag based on the preregistered log-completion-time rule
prolificData$logCompleTime <- log(prolificData$Time.taken)
prolificData$tooFast <- 1 * (
  prolificData$logCompleTime <
    mean(log(prolificData$Time.taken)) - 2 * sd(log(prolificData$Time.taken))
)

keep_ids <- prolificData$Participant.id[prolificData$tooFast == 0]

choices <- subset(
  choices,
  BotCheck1attempt == 1 &
    Part1CQsPassedOnAttempt <= 2 &
    Part2CQsPassedOnAttempt <= 2 &
    Reliable >= 6 &
    AIUsageInStudy == 0 &
    Delegated >= 0 &
    ProlificID %in% keep_ids
)

subjects  <- subset(subjects,  ProlificID %in% choices$ProlificID)
choicesP1 <- subset(choicesP1, ProlificID %in% choices$ProlificID)

# Final analysis sample (paper: N = 233; EV = 118, Pref = 115)
cat("Final sample size per condition (1 = EV, 0 = Preference-based):\n")
table(subjects$EVAlgorithm)

# Convenience label for the algorithm condition
choices <- choices %>%
  mutate(Algorithm = ifelse(EVAlgorithm == 1, "EV", "Pref"))
```

## Build participant-level Part-1 preference measures

These variables (revealed risk preferences from Part 1, per-task confidence,
and post-task expectations) are merged onto the decision-level `choices` data
for use in the regressions and the appendix coefficient plot.

```{r part1-measures}
# EV-maximising choice in each Part-1 task
choicesP1 <- choicesP1 %>%
  mutate(EVChoice = case_when(
    grepl("^P", ChoiceType) & PBetChosen == 1 ~ 1,
    grepl("^D", ChoiceType) & PBetChosen == 0 ~ 1,
    TRUE ~ 0
  ))

# "True" P-bet choice (less-risky choice when it was NOT EV-maximising)
choicesP1 <- choicesP1 %>%
  mutate(truePChoice = case_when(
    grepl("^D", ChoiceType) & PBetChosen == 1 ~ 1,
    TRUE ~ 0
  ))

# Participant-level Part-1 summaries
p1_summaries <- choicesP1 %>%
  group_by(ProlificID) %>%
  summarise(
    EVChoices    = sum(EVChoice),
    truePChoices = sum(truePChoice),
    PBet_count   = sum(PBetChosen),
    .groups = "drop"
  )

choices <- choices %>%
  left_join(p1_summaries, by = "ProlificID") %>%
  # per-task Part-1 confidence (matched on participant x option set)
  left_join(
    choicesP1 %>% select(ProlificID, OptionsetID, ConfidenceP1 = Confidence),
    by = c("ProlificID", "OptionsetID")
  ) %>%
  # participant-level demographics and post-task expectations
  left_join(
    subjects %>% select(
      ProlificID, Female, Age, Education, HHIncome,
      ProbOfBadComputerChoices, ProbOfAbsurdComputerChoices, ComputerRiskTaking
    ),
    by = "ProlificID"
  )
```

\newpage

# Descriptive overview

## Choice quality and payoffs

Choice quality among non-delegated decisions and the payoff advantage of
delegation (paper, Section "Primary test: first-contact delegation").

```{r descriptives-choicequality}
# Share of non-dominated self-chosen options, by condition
choices %>%
  filter(Part == 2) %>%
  group_by(Algorithm) %>%
  summarise(
    avNonDom = mean(Nondominated[Delegated == 0 & Option > 0], na.rm = TRUE),
    .groups = "drop"
  )

# Average payoff: delegated vs. non-delegated decisions
# (paper: 243.49 points delegated vs. 195.18 points non-delegated, ~25% higher)
choices %>%
  filter(Part == 2) %>%
  group_by(Delegated) %>%
  summarise(meanPayoff = mean(RelevantPayoff, na.rm = TRUE), .groups = "drop")
```

## Round-level delegation profile

```{r round-level-table}
choicesAv <- choices %>%
  filter(Part == 2) %>%
  group_by(Algorithm, WithinPartDecisionNumber) %>%
  summarise(
    avDeleg  = mean(Delegated, na.rm = TRUE),
    avNonDom = mean(Nondominated[Delegated == 0 & Option > 0], na.rm = TRUE),
    delegSE = sd(Delegated, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )
choicesAv
```

## Figure 1 --- Average delegation over decisions

```{r figure1-average-delegation, fig.width=7, fig.height=4.5}
fig1 <- choicesAv %>%
  mutate(Algorithm = factor(Algorithm, levels = c("EV", "Pref"),
                            labels = c("EV", "Preference-based"))) %>%
  ggplot(aes(x = WithinPartDecisionNumber, y = avDeleg,
             color = Algorithm, group = Algorithm)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = avDeleg - 1.96 * delegSE, ymax = avDeleg + 1.96 * delegSE),
                width = 0.2, linewidth = 0.8) +
  scale_x_continuous(breaks = 1:8) +
  scale_y_continuous(breaks = (4:7)/10, limits = c(0.325, 0.775)) +
  scale_color_manual(values = c("EV" = col_ev, "Preference-based" = col_pref)) +
  labs(x = "Decision number (Part 2)",
       y = "Average delegation",
       color = "Algorithm") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")

fig1
ggsave("Figure1_AverageDelegation.pdf", fig1, width = 7, height = 4.5)
```

\newpage

# Primary test 1 --- First-contact delegation (round 1)

Preregistered one-sided Boschloo exact test on round-1 delegation
(paper: 66.9% vs. 55.8%, p = 0.0413).

```{r boschloo-round1}
tab <- table(
  Delegated   = choices$Delegated[choices$Part == 2 & choices$WithinPartDecisionNumber == 1],
  EVAlgorithm = choices$EVAlgorithm[choices$Part == 2 & choices$WithinPartDecisionNumber == 1]
)

# Delegation share by condition (column 2 = EV)
prop.table(tab, margin = 2)

# One-sided Boschloo exact test (H1: EV > Pref)
exact.test(tab, method = "Boschloo", alternative = "greater", to.plot = FALSE)$p.value
```

## Effect size and CI for the round-1 contrast

Risk difference (Wald 95% CI) and odds ratio (Fisher's exact two-sided 95% CI), as
reported in the manuscript's new Table 1 (RD = 11.1pp [-1.3, 23.7]; OR = 1.60 [0.91, 2.84]).

```{r boschloo-round1-effectsize}
p_ev   <- prop.table(tab, margin = 2)["1", "1"]
p_pref <- prop.table(tab, margin = 2)["1", "0"]
n_ev   <- sum(tab[, "1"])
n_pref <- sum(tab[, "0"])

# Risk difference + Wald 95% CI
rd <- p_ev - p_pref
se_rd <- sqrt(p_ev * (1 - p_ev) / n_ev + p_pref * (1 - p_pref) / n_pref)
cat("Risk difference:", round(rd, 4), "\n")
cat("Wald 95% CI:", round(rd - 1.96 * se_rd, 4), "to", round(rd + 1.96 * se_rd, 4), "\n")

# Odds ratio + Fisher's exact two-sided 95% CI
fisher.test(tab)
```

\newpage

# Primary test 2 --- Delegation across repeated decisions

## Participant-level Wilcoxon--Mann--Whitney test

Preregistered one-sided WMW test on participant-average delegation
(paper: W = 6428, p = 0.2425).

```{r wmw-test}
delegAv <- choices %>%
  filter(Part == 2) %>%
  group_by(Algorithm, ProlificID) %>%
  summarise(avDeleg = mean(Delegated, na.rm = TRUE), .groups = "drop") %>%
  mutate(Algorithm = factor(Algorithm, levels = c("Pref", "EV")))

wilcox.test(avDeleg ~ Algorithm, data = delegAv, alternative = "less")
```

## Effect size and CI for the WMW test

Rank-biserial correlation with a bootstrap 95% CI, as reported in the manuscript's new
Table 1 ($r_{rb} \approx 0.05$, bootstrap CI [-0.09, 0.21]).

```{r wmw-effectsize}
rank_biserial <- function(x, y) {
  n1 <- length(x); n2 <- length(y)
  U <- wilcox.test(x, y)$statistic
  1 - (2 * U) / (n1 * n2)
}

x_pref <- delegAv$avDeleg[delegAv$Algorithm == "Pref"]
x_ev   <- delegAv$avDeleg[delegAv$Algorithm == "EV"]

r_rb <- rank_biserial(x_pref, x_ev)
cat("Rank-biserial r:", round(r_rb, 4), "\n")

set.seed(1)
B <- 2000
boot_r <- replicate(B, {
  xb <- sample(x_pref, replace = TRUE)
  yb <- sample(x_ev,   replace = TRUE)
  rank_biserial(xb, yb)
})
cat("Bootstrap 95% CI:", round(quantile(boot_r, 0.025), 4), "to",
    round(quantile(boot_r, 0.975), 4), "\n")
```

## Risk aversion
```{r RA-distribution}
# calculate how often the EV-algorithm chooses p-bets
table(choicesP1$ChoiceType[choicesP1$OptionsetID <= 8] == "P105")
# 50-50; contrast that to...
pBetChoices <- choicesP1 %>%
  filter(OptionsetID <= 8) %>% # among the relevant Option sets
  group_by(ProlificID) %>%
  summarise(mean(PBetChosen)) # how often did the participant choose a p-bet
prop.table(table(pBetChoices > 0.5))

```
for 89.7% of the participants, the prefs-algorithm in expectation will produce fewer zero-outcomes than the EV-algorithm


## Figure 3 --- ECDF of participant-level average delegation

```{r figure3-ecdf, fig.width=7, fig.height=4.5}
fig3 <- delegAv %>%
  mutate(Algorithm = factor(Algorithm, levels = c("EV", "Pref"),
                            labels = c("EV", "Preference-based"))) %>%
  ggplot(aes(x = avDeleg, color = Algorithm)) +
  stat_ecdf(linewidth = 1) +
  scale_color_manual(values = c("EV" = col_ev, "Preference-based" = col_pref)) +
  labs(x = "Participant-level average delegation",
       y = "Cumulative probability",
       color = "Algorithm") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")

fig3
ggsave("Figure3_ECDF.pdf", fig3, width = 7, height = 4.5)
```

## Preregistered regression specification

LPM with decision-set fixed effects and participant-clustered SEs. We first
construct the feedback-history regressor (running share of zero outcomes among
prior delegated choices).

```{r feedback-history}
choices <- choices %>%
  arrange(ProlificID, WithinPartDecisionNumber) %>%
  group_by(ProlificID) %>%
  mutate(
    delegated_round = (Delegated == 1),
    zero_payoff     = (RelevantPayoff == 0 & Delegated == 1),
    cum_delegated   = lag(cumsum(delegated_round), default = 0),
    cum_zero        = lag(cumsum(zero_payoff), default = 0),
    zero_rate       = ifelse(cum_delegated > 0, cum_zero / cum_delegated, NA)
  ) %>%
  ungroup()
```

```{r preregistered-regression}
# Preregistered model: treatment x (period + feedback history)
delRegPreReg <- feols(
  Delegated ~ EVAlgorithm * (WithinPartDecisionNumber + zero_rate) | OptionsetID,
  cluster = ~ProlificID,
  data = subset(choices, Part == 2)
)
summary(delRegPreReg)
```

## Robustness check: logistic regression

Re-estimating the preregistered LPM as a logistic regression with the same fixed
effects and cluster-robust SEs (paper: EVAlgorithm OR = 0.79, $p = 0.467$;
zero-outcome-history OR = 0.19, $p < 0.001$).

```{r logit-robustness}
delRegLogit <- feglm(
  Delegated ~ EVAlgorithm * (WithinPartDecisionNumber + zero_rate) | OptionsetID,
  family = "logit",
  cluster = ~ProlificID,
  data = subset(choices, Part == 2)
)
summary(delRegLogit)

# Odds ratios
exp(coef(delRegLogit))
```

## Supporting regressions

Plain treatment effect (all rounds; excluding round 1; round 1 only). The
round-1-only LPM reproduces the Boschloo result
($\beta = 0.113$, one-sided $p = 0.041$).

```{r supporting-regressions}
# All rounds, treatment only
delRegPlainTrmt <- feols(
  Delegated ~ EVAlgorithm | OptionsetID,
  cluster = ~ProlificID,
  data = subset(choices, Part == 2)
)

# Excluding round 1
delRegPlainTrmtNoP1 <- feols(
  Delegated ~ EVAlgorithm | OptionsetID,
  cluster = ~ProlificID,
  data = subset(choices, Part == 2 & WithinPartDecisionNumber > 1)
)

# Round 1 only (mirrors the Boschloo test)
delRegP1 <- feols(
  Delegated ~ EVAlgorithm | OptionsetID,
  cluster = ~ProlificID,
  data = subset(choices, Part == 2 & WithinPartDecisionNumber == 1)
)

screenreg(
  list(delRegPlainTrmt, delRegPlainTrmtNoP1, delRegP1),
  custom.model.names = c("All rounds", "Rounds 2-8", "Round 1 only"),
  caption = "Treatment effect on delegation (LPM, task FE, clustered SE)"
)
```

\newpage

# Categorical non-delegation

We classify participants as never-, sometimes-, or always-delegators and
compare never-delegation across conditions (paper: 15.7% Pref vs. 3.4% EV).

```{r delegator-types}
choices <- choices %>%
  group_by(ProlificID) %>%
  mutate(sumDelegations = sum(Delegated)) %>%
  ungroup() %>%
  left_join(
    prolificData %>% select(Participant.id, TimeInSeconds = Time.taken),
    by = c("ProlificID" = "Participant.id")
  )

analysis_data <- choices %>%
  group_by(ProlificID) %>%
  summarise(
    EVAlgorithm                 = first(EVAlgorithm),
    sumDelegations              = first(sumDelegations),
    Female                      = first(Female),
    Age                         = first(Age),
    HHIncome                    = first(HHIncome),
    Education                   = first(Education),
    ProbOfBadComputerChoices    = first(ProbOfBadComputerChoices),
    ProbOfAbsurdComputerChoices = first(ProbOfAbsurdComputerChoices),
    ComputerRiskTaking          = first(ComputerRiskTaking),
    PBetChoices                 = first(PBet_count),
    TimeInSeconds               = first(TimeInSeconds),
    .groups = "drop"
  ) %>%
  mutate(
    DelegationGroup = ifelse(sumDelegations == 0, "Never", "AtLeastOnce"),
    NonDelegator    = 1 * (sumDelegations == 0),
    DelegatorType   = factor(
      if_else(sumDelegations == 0, "NeverDelegator",
              if_else(sumDelegations == 8, "AlwaysDelegator", "SometimesDelegator")),
      levels = c("NeverDelegator", "SometimesDelegator", "AlwaysDelegator")
    )
  ) %>%
  left_join(
    subjects %>% select(ProlificID, RiskTaking, FinMarketExperience,
                        FinanciallyResponsible, AIXperience),
    by = "ProlificID"
  )

# Distribution of delegator types by condition (column 2 = EV)
table(analysis_data$DelegatorType, analysis_data$EVAlgorithm)
prop.table(table(analysis_data$DelegatorType, analysis_data$EVAlgorithm), margin = 2)
```

## Significance test and effect size for categorical non-delegation

Fisher's exact test, risk difference (Wald 95% CI), and odds ratio (Fisher's exact 95%
CI), as newly added to the manuscript (paper: $p = 0.0014$; RD = 12.3pp [4.9, 19.7];
OR = 5.25 [1.65, 22.07]).

```{r nondelegation-significance}
nondeleg_tab <- table(
  NonDelegator = factor(analysis_data$NonDelegator, levels = c(1, 0)),
  EVAlgorithm  = factor(analysis_data$EVAlgorithm,  levels = c(0, 1))
)
nondeleg_tab   # rows: never-delegator yes/no; columns: 0 = Pref, 1 = EV

# Odds ratio of never-delegating, Pref relative to EV (expect OR ~ 5.25)
fisher.test(nondeleg_tab)

p_pref_nd <- prop.table(nondeleg_tab, margin = 2)["1", "0"]
p_ev_nd   <- prop.table(nondeleg_tab, margin = 2)["1", "1"]
n_pref_nd <- sum(nondeleg_tab[, "0"])
n_ev_nd   <- sum(nondeleg_tab[, "1"])
# (row "1" = never-delegator; column "0" = Pref, "1" = EV)

rd_nd <- p_pref_nd - p_ev_nd
se_rd_nd <- sqrt(p_pref_nd * (1 - p_pref_nd) / n_pref_nd + p_ev_nd * (1 - p_ev_nd) / n_ev_nd)
cat("Risk difference:", round(rd_nd, 4), "\n")
cat("Wald 95% CI:", round(rd_nd - 1.96 * se_rd_nd, 4), "to",
    round(rd_nd + 1.96 * se_rd_nd, 4), "\n")
```

## Comparisons of never-delegators vs. delegators

The following tests reproduce the in-text comparisons (age, income, education,
time, gender, risk-taking, reliability, AI experience, Part-1 satisfaction).

```{r nondelegator-tests}
# Age
t.test(Age ~ DelegationGroup, data = analysis_data)

# Time in experiment
t.test(TimeInSeconds ~ NonDelegator, data = analysis_data)
wilcox.test(TimeInSeconds ~ NonDelegator, data = analysis_data)
ks.test(TimeInSeconds ~ NonDelegator, data = analysis_data)

# Household income (categorical)
chisq.test(table(analysis_data$DelegationGroup, analysis_data$HHIncome))

# Education (categorical)
chisq.test(table(analysis_data$DelegationGroup, analysis_data$Education))

# Gender: never-delegation rate among males vs. females (Boschloo)
female_table <- table(
  DelegationGroup = analysis_data$DelegationGroup,
  Female          = analysis_data$Female
)
exact.test(female_table[, c(2, 4)], method = "Boschloo", to.plot = FALSE)$p.value

# Self-reported financial risk-taking, reliability, AI experience
t.test(RiskTaking  ~ NonDelegator, data = analysis_data)
t.test(FinanciallyResponsible ~ NonDelegator, data = analysis_data)
t.test(AIXperience ~ NonDelegator, data = analysis_data)
```

### Non-parametric robustness check (Wilcoxon) for the comparisons above

As newly added to the manuscript footnote (paper: age $p=0.649$; time $p=0.105$;
risk-taking $p=0.854$; reliability $p=0.956$; AI experience $p=0.445$). Note: the
manuscript's "reliability" item corresponds to `Reliable` in `subjects`/`choices`,
merged in here for completeness; if not already present in `analysis_data`, join it
from `subjects` first.

```{r nondelegator-tests-nonparametric}
wilcox.test(Age ~ NonDelegator, data = analysis_data)
wilcox.test(TimeInSeconds ~ NonDelegator, data = analysis_data)
wilcox.test(RiskTaking ~ NonDelegator, data = analysis_data)
wilcox.test(AIXperience ~ NonDelegator, data = analysis_data)

# Reliability (Reliable), if not yet merged into analysis_data:
if (!"Reliable" %in% names(analysis_data)) {
  analysis_data <- analysis_data %>%
    left_join(subjects %>% select(ProlificID, Reliable), by = "ProlificID")
}
t.test(Reliable ~ NonDelegator, data = analysis_data)
wilcox.test(Reliable ~ NonDelegator, data = analysis_data)
```

## Choice quality among never-delegators (new Limitations check)

Tests whether never-delegators' own non-dominated choice rate (among their self-made
Part-2 decisions) differs from that of participants who delegated at least once, used in
the manuscript to argue against a task-difficulty confound for categorical
non-delegation (paper: 34.7% vs. 33.6%; Welch $t = -0.25$, $p = 0.806$; Wilcoxon
$p = 0.669$).

```{r choicequality-by-nondelegator}
selfchoice <- choices %>%
  filter(Part == 2, Delegated == 0, Option > 0) %>%
  left_join(
    analysis_data %>% select(ProlificID, DelegatorType),
    by = "ProlificID"
  )

# Decision-level check, by delegator type
selfchoice %>%
  group_by(DelegatorType) %>%
  summarise(avNonDom = mean(Nondominated, na.rm = TRUE), n = n(), .groups = "drop")

# Participant-level average non-dominated rate, Never vs. AtLeastOnce
partavg <- selfchoice %>%
  group_by(ProlificID) %>%
  summarise(Nondominated = mean(Nondominated, na.rm = TRUE), .groups = "drop") %>%
  left_join(analysis_data %>% select(ProlificID, DelegationGroup), by = "ProlificID")

partavg %>%
  group_by(DelegationGroup) %>%
  summarise(Nondominated = mean(Nondominated, na.rm = TRUE), n = n(), .groups = "drop")

t.test(Nondominated ~ DelegationGroup, data = partavg)
wilcox.test(Nondominated ~ DelegationGroup, data = partavg)
```

```{r nondelegator-satisfaction-pref}
# Within the preference-based condition: Part-1 satisfaction by delegation type
NeverDelegate <- choices %>%
  select(ProlificID, sumDelegations) %>%
  distinct() %>%
  mutate(sumDeleg0 = (sumDelegations == 0))

subjects2 <- subjects %>% left_join(NeverDelegate, by = "ProlificID")

subjects2 %>%
  filter(EVAlgorithm == 0) %>%
  group_by(sumDeleg0) %>%
  summarise(SatisfiedPart1 = mean(SatisfiedPart1, na.rm = TRUE),
            n = n(), .groups = "drop")
```

\newpage

# Responses to feedback (win-stay / lose-shift)

Win-stay/lose-shift model on rounds 2--8 (paper, Section "Responses to
feedback").

```{r win-stay-lose-shift}
choices <- choices %>%
  group_by(ProlificID) %>%
  arrange(WithinPartDecisionNumber, .by_group = TRUE) %>%
  mutate(
    lagDelegated          = lag(Delegated, 1),
    lagRelevantPayoff     = lag(RelevantPayoff, 1),
    lagSatisfactionRating = lag(SatisfactionRating, 1)
  ) %>%
  ungroup()

choices$stay <- 1 * (choices$Delegated == choices$lagDelegated)

# "Prior win": the outcome of the PREVIOUS round was strictly positive.
# (lagWin is defined from the lagged payoff, i.e. the genuine win-stay/lose-shift
# notion of reacting to the previously experienced outcome.)
choices$lagWin <- 1 * (choices$lagRelevantPayoff > 0)

# Pooled model: treatment x prior win
winStayReg <- feols(
  stay ~ EVAlgorithm * lagWin | OptionsetID + WithinPartDecisionNumber,
  cluster = ~ProlificID,
  data = subset(choices, Part == 2 & WithinPartDecisionNumber > 1)
)
summary(winStayReg)
```

Separately by treatment, distinguishing whether the previous round was
delegated (`lagDelegated == 1`, i.e. "the algorithm" produced the outcome) from
a self-made choice. This separates lose-shift after one's own loss from
lose-shift after a delegated (algorithmic) loss.

```{r win-stay-by-treatment}
# EV condition
winStayReg_EV <- feols(
  stay ~ lagDelegated * I(1 - lagWin) | OptionsetID + WithinPartDecisionNumber,
  cluster = ~ProlificID,
  data = subset(choices, EVAlgorithm == 1 & Part == 2 & WithinPartDecisionNumber > 1)
)
summary(winStayReg_EV)

# Preference-based condition
winStayReg_Pref <- feols(
  stay ~ lagDelegated * I(1 - lagWin) | OptionsetID + WithinPartDecisionNumber,
  cluster = ~ProlificID,
  data = subset(choices, EVAlgorithm == 0 & Part == 2 & WithinPartDecisionNumber > 1)
)
summary(winStayReg_Pref)
```

```{r win-stay-rates}
# Average stay rates by condition, prior delegation, and prior win/loss
choices %>%
  filter(Part == 2, WithinPartDecisionNumber > 1) %>%
  group_by(EVAlgorithm, lagDelegated, lagWin) %>%
  summarise(avg_stay_rate = mean(stay, na.rm = TRUE),
            n = n(), .groups = "drop") %>%
  arrange(EVAlgorithm, lagDelegated, lagWin)
```


\newpage

# Exploratory mechanisms

## Coefficient model (interaction with performance expectations)

LPM of individual delegation decisions interacting treatment with participants'
algorithmic performance expectations, controlling for round, per-choice Part-1
confidence, revealed Part-1 risk preferences, and task fixed effects.

```{r mechanism-regression}
delRegWConf <- feols(
  Delegated ~ EVAlgorithm * (WithinPartDecisionNumber + ConfidenceP1 +
    ProbOfAbsurdComputerChoices + ProbOfBadComputerChoices +
    EVChoices + truePChoices) | OptionsetID,
  cluster = ~ProlificID,
  data = subset(choices, Part == 2)
)
summary(delRegWConf)
```

<!-- ## Figure A.2 --- Expected non-preferred and completely weird choices -->

<!-- ```{r figureA2-expectations, fig.width=7, fig.height=4} -->
<!-- figA2 <- subjects %>% -->
<!--   pivot_longer(cols = c(ProbOfBadComputerChoices, ProbOfAbsurdComputerChoices), -->
<!--                names_to = "variable", values_to = "value") %>% -->
<!--   ggplot(aes(x = value, fill = factor(EVAlgorithm))) + -->
<!--   geom_density(alpha = 0.5) + -->
<!--   facet_wrap(~ variable, labeller = labeller(variable = c( -->
<!--     ProbOfBadComputerChoices    = "Non-preferred choices", -->
<!--     ProbOfAbsurdComputerChoices = "Completely weird choices" -->
<!--   ))) + -->
<!--   scale_fill_manual(values = c("0" = "#E66100", "1" = "#5D3A9B"), -->
<!--                     labels = c("0" = "Preference-based", "1" = "EV"), -->
<!--                     name = "Condition") + -->
<!--   labs(x = "Expected number of choices (out of 10)", y = "Density") + -->
<!--   theme_bw() + -->
<!--   theme(legend.position = "bottom") -->

<!-- figA2 -->
<!-- ggsave("ExpectedBadAndAbsurdChoicesByTrmt.pdf", figA2, width = 7, height = 4) -->
<!-- ``` -->

## Treatment comparison of expected algorithmic errors

Formal comparison of expected non-preferred and completely weird choices across
treatments. A LOWER expected number of non-preferred choices under the
preference-based algorithm would indicate that participants believe the
personalized algorithm is doing its job (matching their preferences).

```{r expectations-comparison}
expect_tbl <- subjects %>%
  mutate(Condition = ifelse(EVAlgorithm == 1, "EV", "Preference-based")) %>%
  group_by(Condition) %>%
  summarise(
    n                  = n(),
    mean_nonpreferred  = mean(ProbOfBadComputerChoices, na.rm = TRUE),
    sd_nonpreferred    = sd(ProbOfBadComputerChoices, na.rm = TRUE),
    mean_weird         = mean(ProbOfAbsurdComputerChoices, na.rm = TRUE),
    sd_weird           = sd(ProbOfAbsurdComputerChoices, na.rm = TRUE),
    .groups = "drop"
  )
expect_tbl

# Expected NON-PREFERRED choices: EV vs. Preference-based
t.test(ProbOfBadComputerChoices ~ EVAlgorithm, data = subjects)
wilcox.test(ProbOfBadComputerChoices ~ EVAlgorithm, data = subjects)

# Expected COMPLETELY WEIRD choices: EV vs. Preference-based
t.test(ProbOfAbsurdComputerChoices ~ EVAlgorithm, data = subjects)
wilcox.test(ProbOfAbsurdComputerChoices ~ EVAlgorithm, data = subjects)

# Within the preference-based condition: do never-delegators expect the
# personalized algorithm to make MORE non-preferred choices than delegators?
# (If not, categorical non-delegation is not driven by distrust in the
#  algorithm's ability to capture one's preferences.)
analysis_data %>%
  filter(EVAlgorithm == 0) %>%
  group_by(NonDelegator) %>%
  summarise(
    n                 = n(),
    mean_nonpreferred = mean(ProbOfBadComputerChoices, na.rm = TRUE),
    .groups = "drop"
  )

t.test(ProbOfBadComputerChoices ~ NonDelegator,
       data = subset(analysis_data, EVAlgorithm == 0))
```

\newpage

# Post-task reasons

## Figure 4 --- Reasons by treatment and delegation type

Reason-selection frequencies for the hypothetical additional-task choice, by
treatment (EV vs. preference-based) and delegation type (delegators vs.
never-delegators).

```{r figure4-reasons, fig.width=8, fig.height=4}
subjects2 <- subjects %>% left_join(NeverDelegate, by = "ProlificID")

subjects_long <- subjects2 %>%
  pivot_longer(cols = DontTrust:Other, names_to = "item", values_to = "clicked")

# Group sizes for legend labels
n_labels <- subjects_long %>%
  group_by(sumDeleg0, EVAlgorithm) %>%
  summarise(n = n_distinct(ProlificID), .groups = "drop") %>%
  mutate(
    group = case_when(
      sumDeleg0 == TRUE  & EVAlgorithm == 0 ~ "Prefs--Never-delegators",
      sumDeleg0 == TRUE  & EVAlgorithm == 1 ~ "EV--Never-delegators",
      sumDeleg0 == FALSE & EVAlgorithm == 0 ~ "Prefs--Delegators",
      sumDeleg0 == FALSE & EVAlgorithm == 1 ~ "EV--Delegators"
    ),
    group_label = paste0(group, " (n=", n, ")")
  )

group_colors <- c(
  "Prefs--Delegators"       = "#9dc3e6",
  "Prefs--Never-delegators" = "#1f4e79",
  "EV--Delegators"          = "#f4b183",
  "EV--Never-delegators"    = "#833c00"
)
label_map <- setNames(n_labels$group_label, n_labels$group)

click_rates_both <- subjects_long %>%
  group_by(sumDeleg0, EVAlgorithm, item) %>%
  summarise(rate = mean(clicked, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    group = case_when(
      sumDeleg0 == TRUE  & EVAlgorithm == 0 ~ "Prefs--Never-delegators",
      sumDeleg0 == TRUE  & EVAlgorithm == 1 ~ "EV--Never-delegators",
      sumDeleg0 == FALSE & EVAlgorithm == 0 ~ "Prefs--Delegators",
      sumDeleg0 == FALSE & EVAlgorithm == 1 ~ "EV--Delegators"
    ),
    group = factor(group, levels = names(group_colors))
  )

fig4 <- ggplot(click_rates_both, aes(x = item, y = rate, fill = group)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.7) +
  scale_fill_manual(values = group_colors, labels = label_map,
                    breaks = names(group_colors), name = "Condition") +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "Item", y = "Click rate") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "right")

fig4
ggsave("Figure4_ReasonsByTrmtAndNeverDelegator2.pdf", fig4, width = 8, height = 4, dpi = 300)
```

## Aggregate reason patterns

In-text figures for performance, distrust, and responsibility motives by
delegation type (delegators vs. never-delegators).

```{r reasons-aggregate}
subjects_long %>%
  group_by(sumDeleg0, item) %>%
  summarise(rate = mean(clicked, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = sumDeleg0, values_from = rate) %>%
  rename(Delegators = `FALSE`, NeverDelegators = `TRUE`)
```

## LPM: never-delegation on reasons

```{r lpm-reasons}
reason_vars <- c("DontTrust", "BetterJob", "HowExactly", "MoreInfo",
                 "Responsibility4Choice", "Responsibility4Outcome",
                 "DislikeRelyingOn", "Reactance2Analysis",
                 "PerformanceCuriosity", "Tedious", "Other")

# Bivariate LPMs (one per reason, HC2 SEs)
bivariate_tbl <- do.call(rbind, lapply(reason_vars, function(rv) {
  m  <- lm(as.formula(paste("as.numeric(sumDeleg0) ~", rv)), data = subjects2)
  ct <- coeftest(m, vcov = vcovHC(m, type = "HC2"))
  data.frame(reason = rv,
             coef = round(ct[2, 1], 3), se = round(ct[2, 2], 3),
             t = round(ct[2, 3], 3), p = round(ct[2, 4], 3))
}))
print(bivariate_tbl, row.names = FALSE)

# Joint LPM: full sample + preference-treatment only
f_joint  <- as.formula(paste("as.numeric(sumDeleg0) ~",
                             paste(reason_vars, collapse = " + ")))
lpm_all  <- lm(f_joint, data = subjects2)
lpm_pref <- lm(f_joint, data = subjects2[subjects2$EVAlgorithm == 0, ])

vcov_all  <- vcovHC(lpm_all,  type = "HC2")
vcov_pref <- vcovHC(lpm_pref, type = "HC2")

screenreg(
  list(lpm_all, lpm_pref),
  override.se = list(sqrt(diag(vcov_all)), sqrt(diag(vcov_pref))),
  override.pvalues = list(
    2 * pt(abs(coef(lpm_all)  / sqrt(diag(vcov_all))),
           df = lpm_all$df.residual,  lower.tail = FALSE),
    2 * pt(abs(coef(lpm_pref) / sqrt(diag(vcov_pref))),
           df = lpm_pref$df.residual, lower.tail = FALSE)
  ),
  custom.model.names = c("Full sample", "Pref. treatment"),
  caption = "LPM: never-delegation ~ reasons (HC2 SEs)"
)
```

\newpage

# Exclusion-criterion breakdown (Method section)

Reconstructs the exact breakdown of the 7 excluded participants by criterion, as
reported in the Method section (paper: 1 bot/slider failure, 4 careful-participation
failures [`Reliable < 6`], 3 disclosed-AI-usage cases, 0 comprehension-only failures,
0 incomplete-data cases, 0 log-time exclusions; final $N = 233$, with 1 participant
overlapping careful-participation-failure and AI-disclosure). This chunk re-reads the
raw files independently of the exclusions already applied above (using `_raw` suffixes),
so it is self-contained and can be run at any point in the script.

```{r exclusion-breakdown}
choices_raw      <- read.delim("260206_1528_537_lotteryChoices.xls")
prolificData_raw <- read.csv("prolific_demographic_export_approved.csv")

cat("N recruited (Prolific export):", nrow(prolificData_raw), "\n")

choices_raw <- subset(choices_raw, ProlificID %in% prolificData_raw$Participant.id)

part2_raw <- subset(choices_raw, Part == 2)
partvars <- part2_raw[!duplicated(part2_raw$ProlificID),
                       c("ProlificID", "BotCheck1attempt", "Part1CQsPassedOnAttempt",
                         "Part2CQsPassedOnAttempt", "Reliable", "AIUsageInStudy")]

# Incomplete-data flag: participant has zero valid (Delegated >= 0) Part-2 rows
valid_counts <- aggregate(Delegated ~ ProlificID, data = part2_raw,
                           FUN = function(x) sum(x >= 0))
names(valid_counts)[2] <- "nValid"
partvars <- merge(partvars, valid_counts, by = "ProlificID", all.x = TRUE)
partvars$incomplete <- is.na(partvars$nValid) | partvars$nValid == 0

# Log-completion-time "too fast" rule
prolificData_raw$logCompleTime <- log(prolificData_raw$Time.taken)
prolificData_raw$tooFast <- 1 * (
  prolificData_raw$logCompleTime <
    mean(log(prolificData_raw$Time.taken)) - 2 * sd(log(prolificData_raw$Time.taken))
)
partvars <- merge(partvars, prolificData_raw[, c("Participant.id", "tooFast")],
                   by.x = "ProlificID", by.y = "Participant.id", all.x = TRUE)

partvars$failBot      <- partvars$BotCheck1attempt != 1
partvars$failComp     <- partvars$Part1CQsPassedOnAttempt > 2 |
                          partvars$Part2CQsPassedOnAttempt > 2
partvars$failCareful  <- partvars$Reliable < 6
partvars$disclosedAI  <- partvars$AIUsageInStudy != 0

excluded <- subset(
  partvars,
  failBot | failComp | failCareful | disclosedAI | incomplete | tooFast == 1
)

cat("\nTotal excluded:", nrow(excluded), "\n")
cat("Fail bot/slider check:", sum(excluded$failBot), "\n")
cat("Fail comprehension check only:", sum(excluded$failComp), "\n")
cat("Fail careful-participation check (Reliable < 6):", sum(excluded$failCareful), "\n")
cat("Disclosed AI usage:", sum(excluded$disclosedAI), "\n")
cat("Incomplete data:", sum(excluded$incomplete), "\n")
cat("Too fast (log-time rule):", sum(excluded$tooFast == 1, na.rm = TRUE), "\n")
cat("Overlap (careful-participation AND disclosed AI):",
    sum(excluded$failCareful & excluded$disclosedAI), "\n")
cat("Final analysis N:", nrow(partvars) - nrow(excluded), "\n")
```

\newpage

# Retrospective power analysis (pilot-based, new Limitations paragraph)

Reconstructs the ex-ante power analysis reported as a new Limitations paragraph: using
only the pilot's effect sizes (not the present study's own results, to avoid the
retrospective-power fallacy; Hoenig \& Heisey, 2001), we compute the minimum detectable
effect (MDE) at 80% power for $n = 100$--120/arm, and the power to re-detect an effect of
the pilot's observed magnitude. Pilot data: `pilot_data_complete_iw.csv` (an earlier,
non-preregistered study reported in Suraci's PhD thesis).

```{r pilot-data}
pilot <- read.csv("pilot_data_complete_iw.csv")
pilot <- subset(pilot, returned == 1)
cat("Pilot N (returned == 1):", nrow(pilot), "\n")

# EV treatments = 1,2; preference-based (RDU) treatments = 3,4
pilot$EValgo <- 1 * (pilot$treatment %in% c(1, 2))
cat("\nPilot N by condition (0 = Pref, 1 = EV):\n")
print(table(pilot$EValgo))

# First-contact delegation (randomized-order "1st" decision)
cat("\nFirst-contact delegation rate by condition (paper: 45.8% Pref vs. 67.4% EV):\n")
print(aggregate(decision1st ~ EValgo, data = pilot, mean))

# Average delegation across all 8 (randomized-order) decisions
decisioncols <- c("decision1st", "decision2nd", "decision3rd", "decision4th",
                   "decision5th", "decision6th", "decision7th", "decision8th")
pilot$avgDeleg <- rowMeans(pilot[, decisioncols], na.rm = TRUE)
cat("\nAverage delegation rate by condition (paper: 35.3% Pref vs. 54.5% EV):\n")
print(aggregate(avgDeleg ~ EValgo, data = pilot, mean))
```

## Minimum detectable effect (MDE) at 80% power

A pure design-stage quantity: depends only on $n$, $\alpha$, and a baseline rate, not on
anything observed in the present study (paper: ~16-17pp for first contact,
~10-11pp for average delegation).

```{r power-mde}
p_pref_firstcontact <- aggregate(decision1st ~ EValgo, data = pilot, mean)[1, 2]

for (n in c(100, 120)) {
  pt <- power.prop.test(n = n, p1 = p_pref_firstcontact, p2 = NULL,
                         power = 0.80, sig.level = 0.05, alternative = "one.sided")
  cat("n =", n, " MDE (first-contact-type proportion) =",
      round(pt$p2 - p_pref_firstcontact, 4), "\n")
}
```

## MDE and power for the average-delegation comparison

Approximated with a one-sided two-sample $t$-test using the pilot's pooled
within-condition SD of participant-average delegation (paper: MDE ~10-11pp;
99.7-99.9% power to re-detect the pilot's ~19pp gap).

```{r power-avgdeleg}
sd_pilot    <- sqrt(mean(tapply(pilot$avgDeleg, pilot$EValgo, var)))
delta_pilot <- diff(tapply(pilot$avgDeleg, pilot$EValgo, mean))
cat("Pilot pooled SD:", round(sd_pilot, 4), " pilot gap:", round(delta_pilot, 4), "\n")

for (n in c(100, 120)) {
  m <- power.t.test(n = n, sd = sd_pilot, power = 0.80, sig.level = 0.05,
                    alternative = "one.sided")
  r <- power.t.test(n = n, delta = delta_pilot, sd = sd_pilot, sig.level = 0.05,
                    alternative = "one.sided")
  cat("n =", n, " MDE =", round(m$delta, 4), " power to re-detect pilot gap =",
      round(r$power, 4), "\n")
}
```

## Power to re-detect the pilot's observed effect

Not a retrospective-power computation on the present study's own effect (which would be
circular); this uses only the pilot's pre-existing effect size as input (paper: 93-96% under the normal approximation, 91-95% by exact-test simulation).

```{r power-redetect}
p_ev_firstcontact <- aggregate(decision1st ~ EValgo, data = pilot, mean)[2, 2]

for (n in c(100, 120)) {
  pt <- power.prop.test(n = n, p1 = p_pref_firstcontact, p2 = p_ev_firstcontact,
                         sig.level = 0.05, alternative = "one.sided")
  cat("n =", n, " power to re-detect pilot's first-contact gap =",
      round(pt$power, 4), "\n")
}
```

\newpage

# Appendix task matrices (Part-1 and Part-2 payoff tables)

Regenerates the full numeric Part-1 (12 binary-choice tasks) and Part-2 (8 option sets)
payoff matrices reported in the manuscript's new Appendix, directly from the prepared
task-construction file `FullTaskTableWithExemplaryOrdering.ods`. Requires the `readODS`
package (`install.packages("readODS")` if not already installed).

```{r appendix-tables}
library(readODS)
task_table <- read_ods("FullTaskTableWithExemplaryOrdering.ods", col_names = FALSE)

# Part-1 block (rows 3:26): 12 tasks x 2 options, excluding the OptionsetID == 0
# instruction rows
part1 <- task_table[3:26, ]
names(part1) <- c("OptionsetID", "Option", paste0("S", 1:10), "Part1DecisionNumber")
part1 <- part1[part1$OptionsetID != 0, ]
cat("Part-1 table: ", nrow(part1), "rows (expect 24 = 12 tasks x 2 options)\n")
part1

# Part-2 block (from row 33): keep only option sets actually used
# (Part2DecisionNumber != -99)
part2_tab <- task_table[33:nrow(task_table), ]
names(part2_tab) <- c("OptionsetID", "Option", paste0("S", 1:10),
                       "Part1DecisionNumber", "Part2DecisionNumber")
part2_tab <- subset(part2_tab, !is.na(OptionsetID) & OptionsetID != 0 &
                       Part2DecisionNumber != -99)
cat("Part-2 table: ", nrow(part2_tab), "rows (expect 64 = 8 option sets x 8 options)\n")
part2_tab
```

```{r session-info}
sessionInfo()
```
